---
title: "Using Machine Learning to Predict the Type of Mistakes in Weight Lifting Exercise"
author: "Lynn Huang"
date: "August 8, 2016"
output: html_document
---

**Background**

In the weight Lifting Excise, the data was collected for five fashions of performing Unilateral Dumbbell Biceps Curl.
Class A: exactly according to the specification; Class B: throwing the elbows to the front; Class C: lifting the dumbbell only halfway; Class D: lowering the dumbbell only halfway and Class E: throwing the hips to the front.
Class A is the correct way, Classes B, C, D, E are the mistakes people usually make. 

**Objective**

The objective of this project is to use machine learning algorithm to create a model to Predict the type of mistakes based on data gathered from a user.

**Load R Package and Data**

```{r,results='hide'}
library(caret)
library(rpart)
library(randomForest)
trainingData<-read.csv("pml-training.csv",
                       stringsAsFactors = TRUE)
testingData<-read.csv("pml-testing.csv",
                      stringsAsFactors = TRUE)
```

**Exploring and Preparing Data**

```{r, results='hide'}
str(trainingData)
str(testingData)

```

```{r}
dim(trainingData)
dim(testingData)

```

There are 19622 obs. of 160 variables (features) in the training set; There are 20 obs. of 160 variables (features) in the training set.  The "classe" variable in the training set is the outcome to predict.

Remove the null,the zero- and near zero-variance Predictors

```{r,results='hide'}
trainingData <- trainingData[, colSums(is.na(trainingData)) == 0] 
testingData <- testingData[, colSums(is.na(testingData)) == 0] 
if (length(nearZeroVar(trainingData)) > 0) {
  trainingData <- trainingData[, -nearZeroVar(trainingData)] 
}
if (length(nearZeroVar(testingData)) > 0) {
  testingData <- testingData[, -nearZeroVar(testingData)] 
}

```


```{r}
dim(trainingData)
dim(testingData)

```

Remove the un-related time and window predictors in the data

```{r}
features<-names(trainingData[7:58])
trainingData <- trainingData[,c(features,"classe")]
testingData <- testingData[,c(features,"problem_id")]

```

**Create training set and testing set for model fitting**

In order to use training data for modeling, testing data for validation, the training data will be sub set into training set and testing set in the modeling/predition process.  The testing data will be used for validation purpose.


```{r}
set.seed(123)
inTrain = createDataPartition(trainingData$classe, p = 3/4, list = F)
training = trainingData[inTrain,]
testing = trainingData[-inTrain,]

```

**Model Selection and Cross Validation**

*Using Decision Tree with CART (rpart)*

```{r}

modFitCART <- rpart(classe ~. , data=training)
predictionsCART <- predict(modFitCART, testing,type="class")
confusionMatrix(predictionsCART, testing$classe)

```


The out of sample error is 25.1% (since accuracy is 74.86%)

*Use Random Forests for prediction*

```{r}
modFitRF <- randomForest(classe ~. , data=training)
predictionsRF <- predict(modFitRF, testing, type = "class")
confusionMatrix(predictionsRF, testing$classe)


```

The out of sample error is 0.5% (since accuracy is 99.5%)
So, Random Forests is used to predict the validation data (original test set)

**Use Random Forests Model to Predict the Users’ Data (20 Cases for Validation)**

```{r}
result <- predict(modFitRF, testingData[, -length(names(testingData))])
print(data.frame(problem_d=testingData$problem_id,result))

```


```{r}


```


```{r}


```

Reference: http://groupware.les.inf.puc-rio.br/har#ixzz4Gi0U23Ir

